---
title: "Assignment 2"
author: "C.L. Jerde"
date: "2022-10-11"
output: html_document
---
## Due by midnight October 19th, 2022 
## This assignment covers lab 3
Directions: Set up a new .Rproj with an Rmarkdown document. The data are available on Gauchospace as copr_plovers.csv and copr_user_days.csv in the assignment 2 folder. Your .rmd  and .html file must be uploaded to Gauchospace. Some questions require only to provide the code. These are marked as 'Code Only.' Other questions require output only or discussion. Please provide your answer with an 'A:.'

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

```

### Question 1 (3 pts) Import the two data files. Format the files to be tidy data in long format. Show code and provide a discription of what you did.  

1. copr_plovers.csv is a data file spanning 2009 to 2020 of the number of observed nests and fledged plovers 

2. copr_user_days.csv is a data file of the registered users of Coal Oil Point Reserve from 2014 to 2020.

```{r, results=FALSE, message=FALSE}
# Imports the data
copr_plovers <- read_csv(here("assignments","assignment 2","copr_plovers.csv"))
copr_user_days <- read_csv(here("assignments","assignment 2","copr_user_days.csv"))
```

```{r, results=FALSE}
#Only the copr_plovers data need to be manipulated into long format

copr_plovers_longer <- copr_plovers %>%
  pivot_longer(cols = '2009':'2020',
               names_to = "year",
               values_to ="counts") %>%
  mutate(year=as.numeric(year))
  
```
A: I did nothing to the copr_user_days.csv as this data is in tidy long form.  However I used the pivot_longer function to make a variable named year and variable named counts that record the number of observed nests or fledged chicks. I also changed the year to numeric.  The copr_plover data are now in tidy long form. 

### Question 2a (3 pts) Using long formatted copr_plover data, join the copr_user_days data to it. Show your code and the first 6 lines of the newly formed data file. No discription is requested. 
```{r, message=FALSE}
plover_user_joined <- copr_plovers_longer %>%
  full_join(copr_user_days)
```

```{r, echo=FALSE}
head(plover_user_joined)
```

### Question 2b (1pt) Why are there so many **NA** in the data file? No code needed.  Provide and answer with A:

A: The data on user days has no observations for 2009-2013.  

### Question 3 (3pts) Create your own function to covert temperature from Fahrenheit to Celsius. Show the code. And then covert -40F and 72F to Celcius by showing code. No description is needed. NOTE: There is something special about the converstion of Celcius and Fahrenheit at -40!

```{r}
# Convert Fahrenheit to Celsius
temp_celcius <- function(F){(F-32) * (5/9)}

temp_celcius(-40)
temp_celcius(72)
```

